---
title: "Customer Churn"
author: "Yashdeep"
date: "29/12/2019"
output:
  pdf_document: default
  word_document: default
---

## Setup root directory for R Markdown

```{r setup}
knitr::opts_knit$set(root.dir = 'C:/Users/Hp/Desktop/R Programming')
```

## Environment Setup

```{r}
setwd("C:/Users/Hp/Desktop/R Programming")
getwd()
```

## Invoke Libraries

```{r}
library(readxl)
library(ggplot2)
library(corrplot)
library(caTools)
library(rms)
library(ineq)
library(ROCR)
library(pROC)
library(class)
library(e1071)
```

## Data Import

```{r}
CellphoneChurnData <- read_excel("Cellphone.xlsx")
```

## Exploratory Data Analysis

```{r}
dim(CellphoneChurnData)
names(CellphoneChurnData)
str(CellphoneChurnData)
summary(CellphoneChurnData)
attach(CellphoneChurnData)
```

## Missing Value Identification

```{r}
is.na(colSums(CellphoneChurnData))
sum(is.na(CellphoneChurnData))
```

## Check for Multicollinearity

```{r}
Matrix <- cor(CellphoneChurnData[,-1])
Matrix
corrplot(Matrix, method = "pie" , type = "upper")

Matrix1 <- cor(CellphoneChurnData[,c(-1,-4,-9)])
Matrix1
corrplot(Matrix1)
corrplot(Matrix1, method = "pie" , type = "upper")
```


## Feature Engineering 

```{r}
CellphoneChurnData$Churn <- as.factor(CellphoneChurnData$Churn)
CellphoneChurnData$ContractRenewal <- as.factor(CellphoneChurnData$ContractRenewal)
CellphoneChurnData$DataPlan <- as.factor(CellphoneChurnData$DataPlan)
CellphoneChurnData$CustServCalls <- as.factor(CellphoneChurnData$CustServCalls)
```

## Univariate Analysis

```{r}
boxplot(CellphoneChurnData, main = "Boxplot for Variables in Churn Dataset")

qplot(DataUsage , geom = "density", xlab = "Gigabytes of Monthly Data Usage",
      ylab = "Frequency",main = "Density Chart of Monthly Data Usage") + theme(plot.title = element_text(hjust = 0.5)) 

qplot(MonthlyCharge , geom = "density", xlab = "Monthly Bill",
      ylab = "Frequency",main = "Density Chart of Customers Monthly Bill") + theme(plot.title = element_text(hjust = 0.5))

qplot(AccountWeeks, geom = "bar", xlab = "Account in Weeks",  
      ylab = "Freqeuncy", main = "Number of weeks customer had an active account") +
  theme(plot.title = element_text(hjust = 0.5)) 

qplot(CellphoneChurnData$RoamMins, geom = "bar", xlab = "Average Roaming Minutes",  
      ylab = "Freqeuncy", main = "Visualization of average roaming minutes used by a customer") + theme(plot.title = element_text(hjust = 0.5))


```

## Bivariate Analysis

```{r}
qplot(CellphoneChurnData$CustServCalls, xlab = "Customer Service Calls", ylab = "Count", 
      main  = "Percentage of Churn by Customer Service Calls", fill = CellphoneChurnData$Churn) + theme(plot.title = element_text(hjust = 0.5)) +  labs(fill='Churn')

qplot(CellphoneChurnData$ContractRenewal, xlab = "Contract Renewal", ylab = "Count", 
      main  = "Percentage of Churn by Contract Renewals", fill = CellphoneChurnData$Churn) +   theme(plot.title = element_text(hjust = 0.5)) +  labs(fill='Churn')

qplot(CellphoneChurnData$DataPlan, xlab = "Data Plan", ylab = "Count", 
      main  = "Percentage of Churn in people with or without Data Plan", fill = CellphoneChurnData$Churn) + 
  theme(plot.title = element_text(hjust = 0.5)) +  labs(fill='Churn')

qplot(CellphoneChurnData$Churn,CellphoneChurnData$DataUsage, geom = "boxplot", xlab = "Churn",  ylab = "Data Usage", main = "Boxplot of Churn against Data Usage") + 
  theme(plot.title = element_text(hjust = 0.5)) 

qplot(CellphoneChurnData$Churn, MonthlyCharge, geom = "boxplot", 
      fill = CellphoneChurnData$Churn, xlab = "Churn", ylab = "Monthly Charge",
      main = "Boxplot of Churn Against Monthly Charge") +  theme(plot.title = element_text(hjust = 0.5))

qplot(DayCalls, MonthlyCharge, xlab = "Average Number of Daytime Calls", 
      ylab = "Average Monthly Bill", main  = "Relationship of Daytime calls V/S Monthly Charge") +  theme(plot.title = element_text(hjust = 0.5))

qplot(DayMins, MonthlyCharge, xlab = "Average Number of Daytime Minutes per Month", 
      ylab = "Average Monthly Bill", 
      main  = "Relationship of Daytime calls V/S Daytime Minutes per Month") + 
  theme(plot.title = element_text(hjust = 0.5))

qplot(DataUsage, MonthlyCharge, xlab = "Gigabytes of Monthly Data Usage", 
      ylab = "Average Monthly Bill", 
      main  = "Relationship of Daytime calls V/S Monthly Data Usage") + 
  theme(plot.title = element_text(hjust = 0.5))

```


## Data Partitioning into Trainset and Testset 

```{r}
seed = 1000
set.seed(seed)
SpL <- sample.split(CellphoneChurnData$Churn, SplitRatio = 0.7)
ChurnDataTrainSet <- subset(CellphoneChurnData[,c(-4,-9)], SpL == TRUE)
ChurnDataTestSet <- subset(CellphoneChurnData[,c(-4,-9)], SpL == FALSE)
dim(ChurnDataTrainSet)
dim(ChurnDataTestSet)
```

## Logisitic Regression

```{r}
Log.Reg.First.Model <- glm(Churn~., data = ChurnDataTrainSet, family = "binomial")
summary(Log.Reg.First.Model)
vif(Log.Reg.First.Model)

Logistic.Reg <- glm(Churn~., data = ChurnDataTrainSet[,c(-2,-7)], family = "binomial")
summary(Logistic.Reg)
vif(Logistic.Reg)

ChurnDataTrainSet$Churn <- as.factor(ChurnDataTrainSet$Churn)
levels(ChurnDataTrainSet$Churn) <- c("No","Yes")
```

## Determining a threshold value 

```{r}
plot(Logistic.Reg$fitted.values)
plot(ChurnDataTrainSet$Churn,Logistic.Reg$fitted.values)
```

## Prediction on trainset using the threshold value

```{r}
Predicted.train = predict(Logistic.Reg, data= ChurnDataTrainSet, type="response")
table(ChurnDataTrainSet$Churn, Predicted.train > 0.18)

Predicted.train.logreg <- ifelse(Logistic.Reg$fitted.values > 0.18, "Yes", "No") 
M <- table(ChurnDataTrainSet$Churn,Predicted.train.logreg)
M
```

## Confusion Matrix (Calculation of sensitivity, specificity, classification error rate and accuracy)

```{r}
summary(ChurnDataTrainSet$Churn)
254/338
279/1995
1716/1995
Classification.error.rate.log.reg.train <- (M[1,2]+M[2,1])/nrow(ChurnDataTrainSet)
Classification.error.rate.log.reg.train*100
Accuracy.log.reg.train <- 1 - Classification.error.rate.log.reg.train
Accuracy.log.reg.train*100
```

## Other Model Evaluations (K.S, AUC & Gini)

```{r}
roc(ChurnDataTrainSet$Churn,Logistic.Reg$fitted.values)
plot(roc(ChurnDataTrainSet$Churn,Logistic.Reg$fitted.values))

ROCRpred <- prediction(Predicted.train,ChurnDataTrainSet$Churn)
Perf <- performance(ROCRpred,"tpr","fpr")
plot(Perf)
KS <- max(Perf@y.values[[1]] - Perf@x.values[[1]])
KS

AUC <- performance(ROCRpred,"auc")
AUC <- as.numeric(AUC@y.values)
AUC

Gini <- ineq(Predicted.train, "gini")
Gini 
```

## Prediction on testset using the threshold value

```{r}
ChurnDataTestSet$Churn <- as.factor(ChurnDataTestSet$Churn)
Predicted.test <- predict(Logistic.Reg, newdata = ChurnDataTestSet, type = "response")
M1 <- table(ChurnDataTestSet$Churn, Predicted.test>0.18)
M1
summary(ChurnDataTestSet$Churn)
```

## Confusion Matrix (Calculation of sensitivity, specificity, classification error rate and accuracy)

```{r}
105/145
109/855
746/855
Classification.error.rate.log.reg.test <- (M1[1,2]+M1[2,1])/nrow(ChurnDataTestSet)
Classification.error.rate.log.reg.test*100
Accuracy.log.reg.test <- 1 - Classification.error.rate.log.reg.test
Accuracy.log.reg.test*100
```

## Other Model Evaluations on Testset (K.S, AUC & Gini)

```{r}
ROCRpred1 <- prediction(Predicted.test,ChurnDataTestSet$Churn)
Perf1 <- performance(ROCRpred1,"tpr","fpr")
plot(Perf1)
KS <- max(Perf1@y.values[[1]] - Perf1@x.values[[1]])
KS

AUC1 <- performance(ROCRpred1,"auc")
AUC1 <- as.numeric(AUC1@y.values)
AUC1

Gini1 <- ineq(Predicted.test, "gini")
Gini1 
```

## Scaling for KNN model

```{r}
ScaledTrainData <- scale(ChurnDataTrainSet[,c(-1,-3,-5)])
head(ScaledTrainData)
ScaledTestData <- scale(ChurnDataTestSet[,c(-1,-3,-5)])
head(ScaledTestData)
```

## Applying KNN

```{r}
KNN.Status <- knn(ScaledTrainData, ScaledTestData, ChurnDataTrainSet$Churn, k = 25 )
summary(KNN.Status)
KNN.table <- table(ChurnDataTestSet$Churn, KNN.Status)
```

## Confusion Matrix for KNN

```{r}
summary(ChurnDataTestSet$Churn)
KNN.table
23/145
0/855
855/855
Class.Error.KNN <- (KNN.table[1,2]+KNN.table[2,1])/nrow(ChurnDataTestSet)
Class.Error.KNN*100
Accuracy.KNN <- 1 - Class.Error.KNN
Accuracy.KNN*100
```

## Applying Naive Bayes

```{r}
NB.Status <- naiveBayes(Churn~., data = ChurnDataTrainSet)
NB.Status
```

## Prediction on Trainset and determining threshold value

```{r}
NB.Predict <- predict(NB.Status, type = "raw", newdata = ChurnDataTrainSet)
plot(ChurnDataTrainSet$Churn,NB.Predict[,2])
NB.Table <- table(ChurnDataTrainSet$Churn,NB.Predict[,2] > 0.2)
```

## Naive Bayes Confusion Matrix on Trainset

```{r}
NB.Table
summary(ChurnDataTrainSet$Churn)
245/338
222/1995
1773/1995
Class.Error.NB <- (NB.Table[1,2]+NB.Table[2,1])/nrow(ChurnDataTrainSet)
Class.Error.NB*100
Accuracy.NB <- 1 - Class.Error.NB
Accuracy.NB*100
```

## Prediction on Testset with the determined threshold value

```{r}
NB.Predict.Test <- predict(NB.Status, type = "raw", newdata = ChurnDataTestSet)
NB.Table.Test <- table(ChurnDataTestSet$Churn,NB.Predict.Test[,2] > 0.2) 
```

## Naive Bayes Confusion Matrix on Testset

```{r}
NB.Table.Test
summary(ChurnDataTestSet$Churn)
100/145
83/855
772/855
Class.Error.NB.Test <- (NB.Table.Test[1,2]+NB.Table.Test[2,1])/nrow(ChurnDataTestSet)
Class.Error.NB.Test*100
Accuracy.NB.Test <- 1 - Class.Error.NB.Test
Accuracy.NB.Test*100
```




